### 2013-04-04 latex/old_table.tex
Updated to version submitted to sheridan

\begin{table}[htb]
\begin{center}
\begin{tabular}{|p{8cm}|}
\hline
{\bf Tasks} \\
\hline
Variety of types of contributions \\
Domain specificity \\
Participation is done to `get something done' \\
Participation is done `for fun' \\
Participation is done to exchange knowledge \\
Participation is done to be social \\
Participants are intrinsically motivated \\
Participants are motivated by extrinsic reward \\
\hline
{\bf Users and Roles} \\
\hline
Generality of Audience \\
Participant autonomy \\
Participant anonymity \\
Extent of hierarchical organisation of roles \\
Clear separation of roles/responsibilities among participants \\
\hline
{\bf Participation} \\
\hline
Participants' (geographic) location is used by the service \\
Participation is done via: \\
\hspace{1cm} mobile devices \\
\hspace{1cm} web browsers \\
\hspace{1cm} apps \\
\hspace{1cm} sensors/sensing (location sensing and wearable devices) \\
Participation is done in a mobile context \\
\hline
{\bf Computation, Output and Benefit} \\
\hline
Output solves a computational problem \\
Output requires human insight \\
Output is a subjective appraisal \\
Participants directly benefit from participating \\
Service/Platform derives benefit from participant participation \\
Participants'' friends/social network benefit from participation \\
Participation is for the benefit of a specific person/group other than the participant or platform owner \\
Participation is for the benefit of society/the world at large  \\
\hline
\end{tabular}
\end{center}
\caption{Consolidated constructs of social machines.} \label{table:constructs}
\end{table}